Detecting changes in dispersion in COVID-19 incidence time series using a negative binomial model

### 2023-08-11 noise_covid_v2.Rmd
more changes for text

$$Cov(\hat{\theta_1},\hat{\theta_2}) = 0 $$
The form of the Wald test statistic is:

@@

First, for every county, the model was fit separately to either side of November 26th, 2020 in order to compute values of $\hat{\theta_1}$ and $\hat{\theta_2}$. A test of the two-sided alternative that theta before is different from theta after (different theta after Thanksgiving) was performed.

@@

To ensure that the hypothesis testing framework is valid, a simulation with known (equal) values of theta on either side of Thanksgiving 2020 was run in order to ensure that the distribution of resulting p-values from the test of the two-sided alternative is approximately uniform on the interval from zero to one. In both the validity and power simulations, representative epidemic curves are used. The area under these curves (final outbreak size) is set to be the population size divided by ten.

To implement these constraints on the integral:
we use a normalizing constant of (n/10) * sqrt(2*pi* $\sigma^2$)

@@

@@

![Simulated p-values for the Wald test](./figures/pvals_sim_Wald.pdf)
![Simulated p-values for the Likelihood-Ratio test](./figures/pvals_sim_LRT.pdf)

There appears to be slight Type I error inflation in both tests.

![Mapping the Thanksgiving hypothesis across the US](./figures/map_Wald_US.pdf)

![Scanning for breakpoints using the LRT statistic](./figures/map_diff_theta_US.pdf)

@@

![The LRT surface](./figures/map_diff_theta_US.pdf)

>>>>

$$Cov(\hat{\theta_1}, \hat{\theta_2}) = 0 $$
The form of the Wald test statistic is:

@@

For every county, the model was fit separately to either side of November 26th, 2020 in order to compute values of $\hat{\theta_1}$ and $\hat{\theta_2}$. A test of the two-sided alternative that $\theta$ before is different from theta after (different theta after Thanksgiving) was performed.

@@

To ensure that the hypothesis testing framework is valid, a simulation with known (equal) values of theta on either side of Thanksgiving 2020 was run in order to ensure that the distribution of resulting p-values from the test of the two-sided alternative is approximately uniform on the interval from zero to one. In both the validity and power simulations, Gaussian and uniform epidemic curves (see figure above which demonstrates correct behavior of the test in either case) are used. The area under these curves (final outbreak size) is set to be the population size divided by ten.

To implement these constraints on the integral in the Gaussian case:

we use a normalizing constant of (n/10) * sqrt(2*pi* $\sigma^2$)

@@

![Type I Error Rate and Power - Wald Test](figures/pvals_sim_Wald_table.pdf)

![Wald Test p-values v. Theta Difference](figures/pvals_sim_Wald.pdf)

@@

![Type I Error Rate and Power - LRT Test](figures/pvals_sim_LRT_table.pdf)

![LRT  p-values v. Theta Difference](figures/pvals_sim_LRT.pdf)

## Biogeography of Superspreading Dynamics

![Mapping the Thanksgiving hypothesis across the US](./figures/map_difftheta_US.pdf)

![LRT Scan Figure](./figures/LRT_scan_figure.pdf)

![The LRT surface](./figures/lrt_surface_figure.pdf)

![The Theta Difference Surface](./figures/thetadiff_surface.pdf)

@@
